---
output:
  html_document:
    keep_md: yes
---
# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
### 1. Load the data
```{r load, echo=TRUE}
library(data.table)
df <- read.csv("activity.csv")
```

### 2. Process/transform the data into a format suitable for analysis
```{r transform}
# conversion to data.table
dt <- as.data.table(df) 
dt$date <- as.Date(dt$date, format="%Y-%m-%d")
# omit NA 
dt2 <- na.omit(dt)
```

## What is mean total number of steps taken per day?

### 1. Make a histogram of the total number of steps taken each day
```{r plothistogram}
dt2[, sumofstepsbydate:=sum(steps), by=date]
plot(dt2$date, dt2$sumofstepsbydate, type="h", col="blue", lwd=5, main = "The total number of steps taken each day", xlab="Date", ylab="# of steps")
```

### 2. Calculate and report the mean and median total number of steps taken per day (Ignore the missing values)
```{r caculation}
report <- dt2[, list(mean(steps),median(steps)), by=date]
setnames(report,2:3,c("mean","median"))
report
plot(dt2$date, dt2$sumofstepsbydate, type="h", col="blue", lwd=5, main = "The total number of steps taken each day", xlab="Date", ylab="# of steps")
plot(report$date,report$mean,col="red",type="l", xlab="Date",ylab="average of steps")
points(report$date,report$median,col="red",type="l", lty=2)
legend("topright",
       c(names(report)[2],names(report)[3]),
       lty=c(1,2),
       col=c("red","red") )
```

## What is the average daily activity pattern?

### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r activitypattern}
activitypattern <- dt2[,mean(steps), by=interval]
setnames(activitypattern,2,"mean")
head(activitypattern)
plot( x=activitypattern$interval, y=activitypattern$mean, type="l", xlab="interval", ylab="average number of steps across all days")
```

### 2. Identifier of 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps
```{r interval}
(aa <- activitypattern$interval[which.max(activitypattern$mean)])
plot( x=activitypattern$interval, y=activitypattern$mean, type="l", xlab="interval", ylab="average number of steps across all days")
axis(1,at=aa, col.ticks="red", col.axis="red")
abline(v=aa, col=2, lty=2)
```

### >> So, Answer is `r aa`. 

## Imputing missing values

### 1. The total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
bad <- !complete.cases(dt)
(a <- sum(bad))
```

### >> So, the total number of rows with NAs in the dataset is `r a`.

### 2. Devise a filling strategy 

### >> Use the mean for that 5-minute interval across all the days

### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in based on logic above.
```{r filling}
dt$steps[which(bad)] <- activitypattern[ match( dt$interval[which(bad)] , activitypattern$interval ),]$mean
head(dt)
```

### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r newhistogram}
report2 <- dt[, list(sum(steps),mean(steps),median(steps)), by=date]
setnames(report2,2:4,c("sum","mean","median"))
plot(report2$date, report2$sum, type="h", col="blue", lwd=5, main = "The total number of steps taken each day", xlab="Date", ylab="# of steps")
report2
```

### 4.1 Do these values differ from the estimates from the first part of the assignment? 
```{r compareaverage}
plot(report$date,report$mean,col="red",type="l", xlab="Date",ylab="average of steps")
points(report$date,report$median,col="red",type="l", lty=2)
points(report2$date,report2$mean, col="blue", type="l")
points(report2$date,report2$median, col="blue", type="l",lty=2)
legend("topright",
       c( paste("before:",names(report)[2]),
          paste("before:",names(report)[3]),
          paste("after:",names(report2)[3]),
          paste("after:",names(report2)[4])
       ),
       lty=c(1,2,1,2),
       col=c("red","red","blue","blue") )
```

#### >> Yes, because of filling missing values, there happens some difference in median plot when averaging daily activity pattern that could be seen above chart.  

### 4.2 What is the impact of imputing missing data on the estimates of the total daily number of steps?

#### >> Big difference in median estimates but relatively small difference in mean when averaging "daily activity pattern"" that could be seen above chart.

## Are there differences in activity patterns between weekdays and weekends?

### 1.Create a new factor variable in the dataset with two levels - ¡°weekday¡± and ¡°weekend¡± indicating whether a given date is a weekday or weekend day.
```{r}
head(dt)
## Change locale 
lct <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_TIME","C")
day <- character()
for ( i in 1:length(dt$date) ) { 
        if ( weekdays(dt$date[i]) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday" ) ) 
               day <- c(day,"Weekday")
            else day <- c(day,"Weekend") 
}
newdt <- cbind(dt,day, stringsAsFactors=T)
head(newdt)
## Recover locale 
Sys.setlocale("LC_TIME",lct)
```

### 2. Make a panel plot containing a time series plot:
```{r latticeplot}
library(lattice)
activitypattern2 <- newdt[,mean(steps), by=c("interval","day")]
setnames(activitypattern2,3,"steps")
with( activitypattern2, xyplot(steps~interval|day, layout=c(1,2), xlab="Interval", ylab="Number of steps", type="l"))
```